\documentclass[11pt, a4paper]{article}

% Packages
\usepackage[margin=1in]{geometry}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{tikz}
\usetikzlibrary{arrows.meta, positioning, shapes.geometric, calc}
\usepackage{circuitikz}
\usepackage{siunitx}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{graphicx}

% Custom commands
\newcommand{\vds}{V_{DS}}
\newcommand{\vgs}{V_{GS}}
\newcommand{\ids}{I_{DS}}
\newcommand{\vth}{V_{th}}
\newcommand{\vdc}{V_{DC}}
\newcommand{\ron}{R_{\text{on}}}

% Title and metadata
\title{From Ideal Switch to Nonlinear ODE: \\
A Pedagogical Guide to SiC MOSFET Switching Transients}
\author{T. Mouli}
\date{\today}

\begin{document}

\maketitle

\tableofcontents
\newpage

\section{The Ideal Switch: Starting Point}

\subsection{The DPT Circuit}

We begin with a \textbf{Double Pulse Test (DPT)} circuit—one of the simplest configurations used to characterize semiconductor switches. The circuit consists of:

\begin{itemize}
    \item A DC voltage source $\vdc = 800\,\text{ V}$
    \item A load inductor $L$ with series resistance $R_L$ (parasitic winding resistance)
    \item A switch $Q$ (our MOSFET)
    \item A freewheeling diode $D$ (body diode of the switch, modeled as piecewise linear)
    \item Gate signal $v_{gs}(t)$ that controls the switch
\end{itemize}

\subsubsection{DPT Circuit Diagram}

Figure~\ref{fig:dpt_circuit} shows the DPT topology. The inductor is connected in series with the DC supply, and the MOSFET with its body diode forms the switch element. The gate signal controls the MOSFET on/off state.

\begin{figure}[h]
\centering
\begin{circuitikz}[scale=1.1, american voltages, american currents,
    every node/.style={font=\small}]

    % ── DC Voltage Source: left vertical branch, + at top ─────
    \draw (0,0) to[V, v^<=$V_{DC}$, a=800\,V] (0,5);

    % ── Top rail: DC+ → R_L → L → switch node (SW) ───────────
    \draw (0,5) to[R, l=$R_L$, f>^=$i_L$] (3,5)
                to[cute inductor, l=$L$]  (6,5);

    % ── Bottom rail (ground/DC−) ───────────────────────────────
    \draw (0,0) to[short] (6,0);

    % ── MOSFET Q: bottom switch
    %    drain = switch node (6,5), source = ground (6,0)
    %    nmos placed with drain anchor at (6,5), gate faces left
    \draw (6,5) node[nmos, anchor=drain, xscale=-1] (Q) {};
    \draw (Q.source) to[short] (6,0);
    \node[right=4pt] at (Q.base) {$Q$};

    % ── Body diode D in anti-parallel with Q ──────────────────
    %    Anode at source node (6,0), cathode at drain node (6,5)
    %    Placed left of Q in the same right branch
    \draw (5.0,0) to[diode, l_=$D$] (5.0,5);
    % Tie diode terminals to the main power rails
    \draw (5.0,5) to[short] (6,5);
    \draw (5.0,0) to[short] (6,0);

    % ── Current direction label in top rail ───────────────────
    \node[above] at (1.5,5) {\textcolor{blue}{$\longrightarrow\;i_L$}};

    % ── Gate drive loop ───────────────────────────────────────
    %    Gate → v_gs source → source of Q (ground)
    \draw (Q.gate) to[short]              (3.5, 2.4)
                   to[V, l_=$v_{gs}(t)$] (3.5, 0)
                   to[short]             (6,   0);

    % ── v_Q: drain–source voltage annotation ──────────────────
    \draw[red, thick, <->] (7.2,5) -- (7.2,0)
        node[midway, right, red, font=\normalsize] {$v_Q$};
    \draw[red, dashed, thin] (6,5) -- (7.2,5);
    \draw[red, dashed, thin] (6,0) -- (7.2,0);

    % ── Node labels ───────────────────────────────────────────
    \node[above right] at (6,5) {\footnotesize SW};
    \node[below]       at (3,0) {\footnotesize GND};

\end{circuitikz}
\caption{Double Pulse Test (DPT) circuit topology.
    A DC source $V_{DC} = 800$\,V (left branch) drives current $i_L$
    through the series load $R_L$--$L$ (top rail) into the switch node (SW).
    The MOSFET $Q$ (bottom switch: drain at SW, source at GND) is driven by
    gate signal $v_{gs}(t)$.
    The body diode $D$ is modelled as a piecewise-linear element in
    \textit{anti-parallel} with $Q$ (anode at source, cathode at drain):
    it provides the freewheeling path for $i_L$ whenever $Q$ is off.
    Drain--source voltage $v_Q$ is indicated on the right.}
\label{fig:dpt_circuit}
\end{figure}

\subsection{The Piecewise Linear MOSFET Model}

In this first section, we use a \textbf{piecewise linear model} for the switch $Q$ with realistic turn-on/off behavior:

\subsubsection{MOSFET Turn-On/Off Logic}

\begin{enumerate}
    \item \textbf{ON state:} When $\vgs \geq \vth$ (gate voltage reaches or exceeds threshold voltage):
    \[
        \ids = \frac{\vds}{\ron}, \quad \text{where } \ron \approx 0
    \]
    The switch conducts with near-zero on-resistance. For simplicity, we approximate $\vds \approx 0$ when conducting.

    \item \textbf{OFF state:} When $\vgs < \vth$, the switch is reverse-biased:
    \[
        \ids = 0
    \]
    No current flows through $Q$. The full supply voltage appears across $\vds$ (minus the diode voltage).

    \item \textbf{No capacitance:} In this first section, the switch has no parasitic energy storage (no $C_{DS}$, no $C_{GS}$, no $C_{GD}$). These will be added later.

    \item \textbf{Threshold voltage:} The switch turns ON when $\vgs$ crosses the threshold:
    \[
        \vth \approx 2\text{--}3\,\text{ V} \quad \text{(typical SiC MOSFET)}
    \]
\end{enumerate}

\subsubsection{Body Diode: Piecewise Linear Model (PWL)}

The freewheeling diode $D$ is modeled using the \textbf{piecewise linear (PWL)} model with three components:
\begin{itemize}
    \item An ideal diode (conducts only in forward direction)
    \item A built-in voltage source $V_{bo}$ (forward voltage drop threshold)
    \item A dynamic resistance $r_d$ (diode on-resistance)
\end{itemize}

The PWL diode equations are:

\begin{enumerate}
    \item \textbf{Forward biased} ($i_d > 0$, diode conducts):
    \[
        v_d = V_{bo} + i_d \cdot r_d
    \]
    where $V_{bo}$ is the built-in voltage (e.g., $0.6$--$1.2$ V for SiC) and $r_d$ is the forward resistance.

    \item \textbf{Reverse biased} ($i_d = 0$, diode blocks):
    \[
        i_d = 0, \quad v_d < V_{bo}
    \]
    No current flows. The voltage $v_d$ is determined by the circuit (can be negative).
\end{enumerate}

This PWL model replaces the more complex Shockley exponential equation with a simple piecewise linear relationship that is easy to solve analytically.

\subsection{Circuit Equations with Piecewise Linear Models}

With the piecewise linear MOSFET and diode models, the circuit behavior depends on the switch state. Let us apply Kirchhoff's voltage law (KVL) around the loop containing the inductor, MOSFET, and diode:

\[
    V_{DC} = i_L R_L + v_Q + v_d
\]

where:
\begin{itemize}
    \item $i_L$ is the inductor current
    \item $R_L$ is the series resistance of the inductor
    \item $v_Q$ is the voltage across the MOSFET drain-source terminals
    \item $v_d$ is the voltage across the body diode
\end{itemize}

\subsubsection{Case 1: MOSFET is ON} ($\vgs \geq \vth$)

When the MOSFET conducts, the switch offers near-zero resistance: $\ron \approx 0 \implies v_Q \approx 0$.

The KVL equation becomes:
\[
    V_{DC} = i_L R_L + v_d
\]

\textbf{Sub-case 1a: Positive current through MOSFET} ($i_L > 0$):

When the inductor current is positive and flowing through the MOSFET, the body diode is reverse-biased:
\begin{itemize}
    \item Ideal diode blocks: $i_d = 0$
    \item All current flows through Q: $i_L = i_Q$
    \item Diode voltage: $v_d = V_{DC} - i_L R_L$
\end{itemize}

\textbf{Sub-case 1b: Current attempts to reverse (freewheeling through diode)}:

When the inductor current reaches zero or tries to reverse, the body diode conducts:
\begin{itemize}
    \item Ideal diode conducts: $i_d > 0$
    \item Diode voltage: $v_d = V_{bo} + i_L r_d$
    \item KVL gives: $V_{DC} = i_L R_L + V_{bo} + i_L r_d = i_L(R_L + r_d) + V_{bo}$
    \item Solving: $i_L = \dfrac{V_{DC} - V_{bo}}{R_L + r_d}$
\end{itemize}

\subsubsection{Case 2: MOSFET is OFF} ($\vgs < \vth$)

When the MOSFET is OFF, all current must flow through the body diode:
\begin{itemize}
    \item Ideal diode conducts: $i_d = i_L$
    \item Diode voltage: $v_d = V_{bo} + i_L r_d$
    \item KVL gives: $V_{DC} = i_L(R_L + r_d) + V_{bo}$
    \item Current: $i_L = \dfrac{V_{DC} - V_{bo}}{R_L + r_d}$
\end{itemize}

\subsection{Summary: Piecewise Behavior}

The circuit exhibits three distinct electrical regimes:

\begin{table}[h]
\centering
\small
\begin{tabular}{lll}
\toprule
\textbf{Regime} & \textbf{Condition} & \textbf{Conducting Path} \\
\midrule
ON, forward & $\vgs \geq \vth$, $i_L > 0$ & MOSFET Q conducts \\
ON, freewheeling & $\vgs \geq \vth$, $i_L < 0$ & Diode D conducts \\
OFF & $\vgs < \vth$ & Diode D conducts \\
\bottomrule
\end{tabular}
\caption{Piecewise linear circuit regimes.}
\end{table}

\subsection{What's Still Missing?}

In this piecewise linear model, all equations are algebraic. However, the inductor has \textbf{inertia}:

\[
    v_L = L \frac{di_L}{dt}
\]

The inductor current cannot change instantaneously. The circuit equations above describe equilibrium states, but not how the current changes with time.

\vspace{1em}
\noindent\textbf{Key question for the next section:}

\framebox{
\parbox{0.85\textwidth}{
If the switch turns OFF and the inductor current is flowing, what equation governs the rate of change $\dfrac{di_L}{dt}$? How does the inductor voltage appear in the circuit during switching?
}
}

\vspace{1em}
\noindent\textit{In the next section, we introduce the inductor as a first-order ODE and solve for $i_L(t)$ during switching transients.}

% ============================================================
\section{The Inductor as a First-Order ODE}
% ============================================================

\subsection{Why the Inductor Matters}

In the previous section, the circuit equations were purely algebraic—we simply looked up which regime the switch was in and read off the voltages and currents. But the load inductor $L$ stores magnetic energy and imposes a fundamental constraint:

\[
    v_L(t) = L\,\frac{di_L}{dt}
\]

This single equation transforms every switching event from an instantaneous jump into a \emph{continuous transient}. The inductor current cannot change instantaneously; instead, it ramps up or down at a rate set by the net voltage across $L$.

\subsection{KVL Around the Main Loop}

Apply Kirchhoff's Voltage Law (KVL) around the main power loop. The sign convention is:
\begin{itemize}
    \item $V_{DC}$ is the driving source (positive, left side of loop).
    \item $i_L R_L$ is the resistive drop across the winding resistance.
    \item $v_Q$ is the voltage across the MOSFET drain–source terminals.
    \item $v_d$ is the voltage across the body diode.
    \item $v_L = L\,\dfrac{di_L}{dt}$ is the volt–second balance across the inductor.
\end{itemize}

The complete KVL equation is:
\[
    V_{DC} = i_L R_L + L\,\frac{di_L}{dt} + v_Q + v_d
\]

Rearranging to isolate the derivative:
\[
    \frac{di_L}{dt} = \frac{1}{L}\Bigl(V_{DC} - i_L R_L - v_Q - v_d\Bigr)
\]

This is a \textbf{first-order linear ODE} in $i_L(t)$. The right-hand side changes depending on which switch state (regime) is active.

\subsection{ODE in Each Switching Regime}

\subsubsection{Regime A: MOSFET ON, Diode Blocked}

When $v_{gs} \geq v_{th}$ and $i_L > 0$:
\[
    v_Q = 0, \quad i_d = 0, \quad v_d = \underbrace{V_{DC} - i_L R_L}_{\text{from sub-case 1a above}}
\]

Wait—if the diode is blocked, $v_d$ is \emph{not} fixed by the PWL equation; instead it floats to satisfy KVL. The conducting element is the MOSFET alone:

\[
    V_{DC} = i_L R_L + L\,\frac{di_L}{dt} + 0
\]
\[
    \boxed{\frac{di_L}{dt} = \frac{V_{DC} - i_L R_L}{L}}
\]

This is a first-order linear ODE with the standard form $\dfrac{di_L}{dt} + \dfrac{R_L}{L}\,i_L = \dfrac{V_{DC}}{L}$.

\subsubsection{Regime B: MOSFET OFF, Diode Conducting (Freewheeling)}

When $v_{gs} < v_{th}$ and $i_L > 0$, the inductor forces current through the body diode (the MOSFET is OFF so $i_Q = 0$, $v_Q = V_{DC} - v_d - i_L R_L$ is absorbed by the circuit):

\[
    v_d = V_{bo} + i_L\,r_d
\]

The KVL (with $v_Q$ floating and the diode in the loop) gives:
\[
    0 = i_L R_L + L\,\frac{di_L}{dt} + (V_{bo} + i_L\,r_d)
\]
\[
    \boxed{\frac{di_L}{dt} = -\frac{(R_L + r_d)}{L}\,i_L - \frac{V_{bo}}{L}}
\]

Here $V_{DC}$ does \emph{not} appear because the DC supply is disconnected from the freewheeling path when $Q$ is OFF—the energy circulates through $L$, $R_L$, and $D$ only.

\subsection{Analytical Solutions}

Both ODEs share the same structure:
\[
    \frac{di_L}{dt} + \frac{R_{\text{eff}}}{L}\,i_L = \frac{V_{\text{eff}}}{L}
\]

where $R_{\text{eff}}$ and $V_{\text{eff}}$ depend on the regime. The general solution is:

\[
    i_L(t) = i_{L,\infty} + \bigl(i_{L,0} - i_{L,\infty}\bigr)\,e^{-t/\tau}
\]

with the time constant $\tau = L / R_{\text{eff}}$ and the steady-state asymptote $i_{L,\infty} = V_{\text{eff}} / R_{\text{eff}}$.

\subsubsection{Regime A — Turn-On Transient}

\begin{align*}
    R_{\text{eff}} &= R_L, \quad V_{\text{eff}} = V_{DC}\\
    \tau_{\text{on}} &= \frac{L}{R_L}\\
    i_{L,\infty} &= \frac{V_{DC}}{R_L}
\end{align*}

Starting from initial current $i_{L,0}$ (typically zero for the first pulse):
\[
    i_L(t) = \frac{V_{DC}}{R_L}\Bigl(1 - e^{-t/\tau_{\text{on}}}\Bigr), \qquad 0 \leq t \leq t_1
\]

The current ramps upward toward $V_{DC}/R_L$. Because $R_L$ is very small (milliohms), $\tau_{\text{on}} = L/R_L$ is large (milliseconds), so within a short pulse duration $t_1 \ll \tau_{\text{on}}$ the ramp is nearly linear:
\[
    i_L(t) \approx \frac{V_{DC}}{L}\,t \qquad \text{(linear approximation for } t \ll \tau_{\text{on}}\text{)}
\]

\subsubsection{Regime B — Turn-Off (Freewheeling) Transient}

\begin{align*}
    R_{\text{eff}} &= R_L + r_d, \quad V_{\text{eff}} = -V_{bo}\\
    \tau_{\text{off}} &= \frac{L}{R_L + r_d}\\
    i_{L,\infty} &= -\frac{V_{bo}}{R_L + r_d}
\end{align*}

Starting from the peak current $i_{L,\text{peak}}$ at turn-off:
\[
    i_L(t) = \left(i_{L,\text{peak}} + \frac{V_{bo}}{R_L + r_d}\right)e^{-t/\tau_{\text{off}}} - \frac{V_{bo}}{R_L + r_d}
\]

The current decays exponentially. In practice $i_L$ is forced to zero by the circuit when the second pulse arrives; the steady-state asymptote $-V_{bo}/(R_L+r_d)$ is never reached.

\subsection{Numerical Example}

Consider the following DPT parameters:

\begin{table}[h]
\centering
\begin{tabular}{lll}
\toprule
\textbf{Parameter} & \textbf{Symbol} & \textbf{Value} \\
\midrule
DC bus voltage      & $V_{DC}$      & 800 V  \\
Load inductance     & $L$           & 500 $\mu$H \\
Winding resistance  & $R_L$         & 10 m$\Omega$ \\
Diode forward drop  & $V_{bo}$      & 1.2 V \\
Diode resistance    & $r_d$         & 5 m$\Omega$ \\
First pulse width   & $t_1$         & 10 $\mu$s \\
\bottomrule
\end{tabular}
\caption{Numerical parameters for the first-order ODE example.}
\label{tab:ode_params}
\end{table}

\textbf{Step 1: Time constants.}
\[
    \tau_{\text{on}} = \frac{500\times10^{-6}}{10\times10^{-3}} = 50\,\text{ms}, \qquad
    \tau_{\text{off}} = \frac{500\times10^{-6}}{15\times10^{-3}} \approx 33.3\,\text{ms}
\]

Since $t_1 = 10\,\mu\text{s} \ll \tau_{\text{on}}$, the linear approximation is valid.

\textbf{Step 2: Peak current after first pulse.}
\[
    i_{L,\text{peak}} \approx \frac{V_{DC}}{L}\,t_1 = \frac{800}{500\times10^{-6}}\times10\times10^{-6} = 16\,\text{A}
\]

\textbf{Step 3: Freewheeling decay.}

After turn-off, the current decays from 16 A. The time for it to reach, say, half its value:
\[
    t_{1/2} = \tau_{\text{off}}\ln 2 \approx 33.3\,\text{ms} \times 0.693 \approx 23\,\text{ms}
\]

This is much longer than typical pulse periods, confirming that $i_L$ barely decays between pulses in a DPT—exactly why two pulses are needed to pre-charge the inductor before the main switching event.

\subsection{What the ODE Captures — and What It Misses}

\begin{table}[h]
\centering
\small
\begin{tabular}{p{4.5cm}p{4.5cm}p{4.5cm}}
\toprule
\textbf{Captured} & \textbf{Assumed} & \textbf{Missing} \\
\midrule
Exponential ramp of $i_L$ & Ideal MOSFET ($v_Q = 0$ when ON) & Gate charge and $V_{GS}$ dynamics \\
Freewheeling current decay & PWL diode model & Parasitic drain–source capacitance $C_{DS}$ \\
Time constant $\tau = L/R$ & No reverse recovery & Gate–drain coupling ($C_{GD}$) \\
Peak current calculation & Zero switching time & Voltage overshoot on $v_{DS}$ \\
\bottomrule
\end{tabular}
\caption{Summary of what the first-order inductor ODE captures and what remains to be modelled.}
\label{tab:ode_summary}
\end{table}

\noindent In particular, notice that \emph{when} the switch turns ON or OFF, the current changes happen with no voltage transient: $v_{DS}$ snaps from $V_{DC}$ to $0$ (or vice versa) instantaneously. Real SiC MOSFETs exhibit \textbf{voltage and current overshoot} during switching, arising from parasitic capacitances. These effects appear only when we add $C_{DS}$, $C_{GD}$, and gate dynamics in later sections.

\subsection{Key Takeaway}

\begin{center}
\fbox{\parbox{0.8\textwidth}{%
\textbf{The inductor turns every switching event into a differential equation.}\\
Its current evolves in time as $i_L(t) = i_{L,\infty} + (i_{L,0} - i_{L,\infty})\,e^{-t/\tau}$,
with time constant $\tau = L/R_{\text{eff}}$. As long as the switch models are ideal
(no capacitances, no gate delay), this ODE has a closed-form analytical solution.
}}
\end{center}

\vspace{1em}
\noindent\textbf{Key question for the next section:}

\framebox{
\parbox{0.85\textwidth}{
In the freewheeling turn-off transient, we assumed the body diode turns on \emph{instantaneously} and carries all of $i_L$. But real diodes have a \textbf{reverse-recovery charge} $Q_{rr}$: before a diode can block, it must first expel stored minority carriers, creating a brief negative current spike. How does this modify the turn-on waveform of the MOSFET---and what does it imply for switching losses?
}
}

\vspace{1em}
\noindent\textit{In the next section, we model the body diode reverse recovery using a charge-based approach and incorporate it into the inductor ODE to capture the current overshoot at turn-on.}

\end{document}
