% Section 1: The Ideal Switch with Realistic Components

\subsection{The DPT Circuit}

We begin with a \textbf{Double Pulse Test (DPT)} circuit—one of the simplest configurations used to characterize semiconductor switches. The circuit consists of:

\begin{itemize}
    \item A DC voltage source $\vdc = 800\,\si{\volt}$
    \item A load inductor $L$ with series resistance $R_L$ (parasitic winding resistance)
    \item A switch $Q$ (our MOSFET)
    \item A freewheeling diode $D$ (body diode of the switch, modeled as piecewise linear)
    \item Gate signal $\vgs(t)$ that controls the switch
\end{itemize}

\noindent The circuit diagram is shown in Figure~\ref{fig:dpt_circuit}.

\begin{figure}[h]
    \centering
    \begin{tikzpicture}[
        scale=1.2,
        thick
    ]
        % Supply
        \node[draw, circle, minimum size=0.8cm] (vsrc) at (0,4) {$+$};
        \node at (-0.15, 3.7) {$V_{DC}$};

        % Load inductor
        \node[draw] (L) at (1.5, 4) [rectangle, minimum width=0.8cm, minimum height=0.5cm] {L};

        % Current arrow
        \node[below=0.3cm of L] (Ilabel) {$I_L$};

        % Diode
        \node[draw] (D) at (3.5, 4) [rectangle, minimum width=0.8cm, minimum height=0.5cm] {D};

        % Switch Q
        \node[draw] (Q) at (3.5, 2) [rectangle, minimum width=0.8cm, minimum height=0.5cm] {Q};

        % Voltages
        \draw[->] (3.8, 3.3) -- (3.8, 2.8) node[right, font=\small] {$v_{ds}$};
        \draw[->] (3.2, 1.5) -- (3.2, 0.8) node[left, font=\small] {$i_d$};

        % Ground symbol
        \draw[-] (0, 0.2) -- (0, 0);
        \draw[-] (-0.2, 0) -- (0.2, 0);
        \draw[-] (-0.15, -0.1) -- (0.15, -0.1);
        \draw[-] (-0.1, -0.2) -- (0.1, -0.2);

        % Gate signal
        \draw (-1, 0.5) -- (1.5, 0.5) node[above, font=\small] {$v_{gs}(t)$};

        % Connections
        \draw (vsrc) -- (L);
        \draw (L) -- (D);
        \draw (D) -- (Q);
        \draw (Q) -- (0, 2);
        \draw (0, 2) -- (0, 0);
        \draw (3.5, 4) -- (3.5, 2.8);
        \draw (3.5, 1.2) -- (3.5, 0);
        \draw (3.5, 0) -- (0, 0);

    \end{tikzpicture}
    \caption{Double Pulse Test (DPT) circuit: DC source + load inductor + MOSFET + body diode.}
    \label{fig:dpt_circuit}
\end{figure}

\subsection{The Piecewise Linear MOSFET Model}

In this first section, we use a \textbf{piecewise linear model} for the switch $Q$ with realistic turn-on/off behavior:

\subsubsection{MOSFET Turn-On/Off Logic}

\begin{enumerate}
    \item \textbf{ON state:} When $\vgs \geq \vth$ (gate voltage reaches or exceeds threshold voltage):
    \[
        \ids = \frac{\vds}{\ron}, \quad \text{where } \ron \approx 0
    \]
    The switch conducts with near-zero ``on-resistance''. For simplicity, we approximate $\vds \approx 0$ when conducting.

    \item \textbf{OFF state:} When $\vgs < \vth$, the switch is reverse-biased:
    \[
        \ids = 0
    \]
    No current flows through $Q$. The full supply voltage appears across $\vds$ (minus the diode voltage).

    \item \textbf{No capacitance:} In this first section, the switch has no parasitic energy storage (no $C_{DS}$, no $C_{GS}$, no $C_{GD}$). These will be added later.

    \item \textbf{Threshold voltage:} The switch turns ON when $\vgs$ crosses the threshold:
    \[
        \vth \approx 2\text{--}3\,\si{\volt} \quad \text{(typical SiC MOSFET)}
    \]
\end{enumerate}

\subsubsection{Body Diode: Piecewise Linear Model (PWL)}

The freewheeling diode $D$ is modeled using the \textbf{piecewise linear (PWL)} model with three components:
\begin{itemize}
    \item An ideal diode (conducts only in forward direction)
    \item A built-in voltage source $V_{bo}$ (forward voltage drop threshold)
    \item A dynamic resistance $r_d$ (diode on-resistance)
\end{itemize}

The PWL diode equations are:

\begin{enumerate}
    \item \textbf{Forward biased} ($i_d > 0$, diode conducts):
    \[
        \vds = V_{bo} + i_d \, r_d
    \]
    where $V_{bo}$ is the built-in voltage (e.g., $0.6\text{--}1.2\,\si{\volt}$ for SiC) and $r_d$ is the forward resistance.

    \item \textbf{Reverse biased} ($i_d = 0$, diode blocks):
    \[
        i_d = 0, \quad \vds < V_{bo}
    \]
    No current flows. The voltage $\vds$ is determined by the circuit (can be negative).
\end{enumerate}

This PWL model replaces the more complex Shockley exponential equation with a simple piecewise linear relationship that is easy to solve analytically.

\noindent\textbf{Key insight:} Once we replace the junction diode with this PWL equivalent, the circuit becomes purely resistive and algebraic—no implicit equations needed.

\subsection{Circuit Equations with Piecewise Linear Models}

With the piecewise linear MOSFET and diode models, the circuit behavior depends on the switch state. Let us apply Kirchhoff's voltage law (KVL) around the loop containing the inductor, MOSFET, and diode:

\[
    \vdc = i_L R_L + v_Q + v_D
\]

where:
\begin{itemize}
    \item $i_L$ is the inductor current (positive flowing from source through Q or D back to ground)
    \item $R_L$ is the series resistance of the inductor (parasitic winding resistance)
    \item $v_Q$ is the voltage across the MOSFET drain-source terminals
    \item $v_D$ is the voltage across the body diode
\end{itemize}

The inductor current $i_L$ flows through one or both devices depending on the switch state.

\subsubsection{Case 1: MOSFET is ON} ($\vgs \geq \vth$)

When the MOSFET conducts, the switch offers near-zero resistance: $\ron \approx 0 \implies v_Q \approx 0$.

The KVL equation becomes:
\[
    \vdc = i_L R_L + v_D
\]

\textbf{Sub-case 1a: Positive current through MOSFET} ($i_L > 0$):

When the inductor current is positive and flowing through the MOSFET, the body diode is reverse-biased and does not conduct:
\begin{itemize}
    \item Ideal diode blocks: $i_d = 0$
    \item Diode voltage constraint: $v_D \leq V_{bo}$ (no current can flow backward)
    \item All current flows through Q: $i_L = i_Q$
    \item Since there is no current through the diode branch, and the MOSFET is ON with $v_Q = 0$:
    \[
        v_D = \vdc - i_L R_L
    \]
\end{itemize}

For the diode to remain reverse-biased, we need:
\[
    v_D = \vdc - i_L R_L < V_{bo}
\]

This condition is typically satisfied when $i_L$ is small.

\textbf{Sub-case 1b: Current attempts to reverse (freewheeling through diode)} ($i_L \to 0^-$):

When the inductor current reaches zero or tries to reverse direction, the body diode conducts:
\begin{itemize}
    \item Ideal diode conducts: $i_d > 0$
    \item Diode voltage: $v_D = V_{bo} + i_d r_d = V_{bo} + i_L r_d$ (since all current flows through diode: $i_L = i_d$)
    \item MOSFET voltage: $v_Q = 0$ (still ON)
    \item KVL gives:
    \[
        \vdc = i_L R_L + V_{bo} + i_L r_d = i_L(R_L + r_d) + V_{bo}
    \]
    \item Solving for inductor current:
    \[
        i_L = \frac{\vdc - V_{bo}}{R_L + r_d}
    \]
\end{itemize}

This is an algebraic equation. The inductor current reaches this value when the diode begins to conduct (at the sub-case transition).

\subsubsection{Case 2: MOSFET is OFF} ($\vgs < \vth$)

When the MOSFET is OFF, the gate voltage is below threshold and the switch blocks: $i_Q = 0$ (no current can flow through Q).

All current must commutate to the body diode:
\begin{itemize}
    \item Ideal diode conducts: $i_d = i_L$ (all current flows through diode)
    \item Diode voltage: $v_D = V_{bo} + i_L r_d$
    \item MOSFET voltage: $v_Q = \vdc - i_L R_L - v_D$
    \[
        v_Q = \vdc - i_L R_L - (V_{bo} + i_L r_d) = \vdc - V_{bo} - i_L(R_L + r_d)
    \]
\end{itemize}

For positive current flowing through the diode:
\[
    i_L = \frac{\vdc - V_{bo}}{R_L + r_d} > 0
\]

The MOSFET sees a voltage:
\[
    v_Q = \vdc - V_{bo} - (R_L + r_d) \cdot \frac{\vdc - V_{bo}}{R_L + r_d} = 0
\]

Wait, that does not look right. Let me reconsider. When the MOSFET is OFF and the diode conducts with current $i_L > 0$:

\[
    \vdc = i_L(R_L + r_d) + V_{bo}
\]

This uniquely determines $i_L$ when given $V_{bo}$, $R_L$, $r_d$, and $\vdc$. However, $i_L$ cannot change instantaneously (due to the inductor), so this algebraic equation describes the \textbf{equilibrium current} that the inductor \textit{approaches}.

\subsection{Summary: Piecewise Behavior}

The circuit exhibits three distinct electrical regimes:

\begin{table}[h]
\centering
\small
\begin{tabular}{llll}
\toprule
\textbf{Regime} & \textbf{Condition} & \textbf{Conducting Path} & \textbf{Key Equation} \\
\midrule
ON, forward & $\vgs \geq \vth, \, i_L > 0$ & MOSFET Q conducts & $v_D = \vdc - i_L R_L$ \\
ON, freewheeling & $\vgs \geq \vth, \, i_L < 0$ & Diode D conducts & $i_L = \dfrac{\vdc - V_{bo}}{R_L + r_d}$ \\
OFF & $\vgs < \vth$ & Diode D conducts & $i_L = \dfrac{\vdc - V_{bo}}{R_L + r_d}$ \\
\bottomrule
\end{tabular}
\caption{Piecewise linear circuit regimes and equations.}
\label{tab:pwl_regimes}
\end{table}

\subsection{What's Still Missing?}

In this piecewise linear model, all equations are \textbf{algebraic} at the equilibrium state. However, the inductor acts like a \textbf{current source with inertia}:

\[
    v_L = L \frac{di_L}{dt}
\]

The inductor current cannot change instantaneously, no matter how large the applied voltage. The circuit equations above tell us what \textit{equilibrium} current each regime would prefer, but not how fast the current \textit{changes} to reach that equilibrium.

In the next section, we will introduce $L$ as a \textbf{state variable} and solve the resulting first-order differential equation.

\vspace{1em}
\noindent\textbf{Key question for the next section:}
\begin{center}
    \fbox{\parbox{0.85\textwidth}{
        If the switch turns OFF and the inductor current is flowing, what equation governs the rate of change $\dfrac{di_L}{dt}$?
        How does the inductor voltage $v_L = L \dfrac{di_L}{dt}$ appear in the circuit during commutation?
    }}
\end{center}

\vspace{1em}
\noindent\textit{In the next section, we introduce the inductor as a first-order ODE and solve for $i_L(t)$ during switching transients.}
